Presentation Title

### 2017-08-06 thesis/chapters/ch5-implementation.tex
Write preface and update Bitcoin situation
Spelling/grammar check for ch3, 4, 5, 6

To build confidence of the horizontal scalability claim from our analytical results,
we must verify the same property experimentally.
Hence, we give an open source prototype implementation of our system.
Our implementation runs on a network with over 1000 node.
Each node send transactions to each other autonomously, 
which allows us to evaluate our analytical claims.

@@

The structure of the implementation is primarily made up of three modules---\texttt{acs}, \texttt{trustchain} and \texttt{node}.
\texttt{acs}, as its name suggests, implements ACS.
ACS uses erasure code in one of its subprotocols (reliable broadcast).
Thus we use the liberasurecode\footnote{\url{https://github.com/openstack/liberasurecode}} library for its Reed-Solomon error correcting code functionality.
An implementation detail is that liberasurecode cannot create more than 32 code blocks\footnote{
  The 32 code blocks limitation is hardcoded in the source file, see
  \url{https://github.com/openstack/liberasurecode/blob/0794b31c623e4cede76d66be730719d24debcca9/include/erasurecode/erasurecode.h\#L35}
}, we discuss the effect of this in~\Cref{sec:experimental-setup}.
The \texttt{acs} module provides a small interface to the caller to start and stop the consensus process and also retrieve results.
The \texttt{trustchain} module implements the Extended TrustChain data structure.
It also provides the essential algorithm necessary to interact with Extended TrustChain such as 
$\textsf{new\_tx}(\cdot)$, $\textsf{new\_cp}(\cdot)$ and $\textsf{agreed\_fragment}(\cdot)$.
Finally, the \texttt{node} module ties everything together.
It implements the the consensus protocol, the transaction protocol and the validation protocol.

@@

The the goal of the experiment is to run the three protocols---consensus protocol,
transaction protocol and validation protocol---simultaneously and analyse the communication cost, the consensus duration and the global throughput.
We investigate these properties under the following four parameters.
\begin{enumerate}
  \item The transaction rate $r_{\text{tx}}$ per node. This is not comparable to the others because it is fixed at 2 TX/s.
        Nevertheless, it is a good value because, as we show later,
        it hits a bottleneck in extreme cases which helps us understand the limitations of our design.
  \item The number of facilitators $n \in \{4, 8, \dots, 32\}$.
        The maximum $n$ is 32 because the limitation in liberasurecode mentioned in \Cref{sec:implementation},
        but our results give a good dindication of how our system may function for larger numbers of $n$.
  \item The population size $N \in \{200, 300, \dots, 1200\}$.
        $N$ stops at 1200 is due to our physical setup, which we describe next.
  \item The two modes of transaction.
        The first mode is the \emph{fixed-neighbour} mode where nodes only transact with their immediate neighbour which are predetermined. 
        This minimises the data volume per validated transaction because agreed fragment can be cached.
        The second mode is in the other extreme which we call the \emph{random-neighbour} mode,
        where every transaction is with a random node out of the $N$ nodes in the system.
        It is unlikely that the agreed fragments can be reused.
\end{enumerate}

The experiment is run on the DAS-5\footnote{\url{https://www.cs.vu.nl/das5/}}.
From now on, we use ``machines'' to refer to DAS-5 nodes and ``nodes'' to refer to a running instance in our system.
On DAS-5 we use up to 30 machine, for each machine we use up to 40 nodes.
This gives us the aforementioned 1200 number.
With this setup, we cannot run more nodes because the every machine only has 65535 ports available (minus the reserved ones).
But 40 nodes each need 1200 TCP connections which is 48000 TCP connections per machine and that is inching close to the limit.
While it is possible to have many more TCP connections per machine,
but it requires additional network interface which is something we do not control on the DAS-5.
Nevertheless, running the system with 1200 nodes gives a good indication of its scalability properties as we show later.

To coordinate nodes on many different machine,
we employ a discovery server to inform every node the IP addresses and port numbers of every other node.
It is only run before the experiment and is not used during the experiment.

@@

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
    \begin{subfigure}[b]{.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-fixed/round-duration-vs-population}
      \caption{Transactions are with fixed neighbours.}
      \label{fig:round-duration-fixed}
    \end{subfigure}%
    \begin{subfigure}[b]{0.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-random/round-duration-vs-population}
      \caption{Transactions are with random neighbours.}
      \label{fig:round-duration-random}
    \end{subfigure}%
  }
  \caption{Round duration do not increase linearly with the population size. 
  This is likely due to the additional hashing operation required for larger consensus result.}
  \label{fig:round-duration}
\end{figure}

\Cref{fig:round-comms} show the relationship between the communication cost of the consensus protocol per round and the population size.
The most important aspect is that these results show a linear increase.
This reinforces our analytical result in~\Cref{sec:cons-complexity}.
Note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
the magnitude of the communication cost are similar.
Both peak at about 100 MB.
This is expected because the consensus protocol is decoupled from the transaction protocol and the validation protocol.
Finally, the rate for which the communication cost increases is higher when the number of facilitators are higher.
This is also expected because the ACS algorithm has a $n^2$ term in its communication complexity.

We are also interested in how communication costs translate to time.
Hence, for the same experiment we record the duration in seconds and the result is shown in~\Cref{fig:round-duration}.
Interestingly, the duration is not entirely linear.
We attribute this behaviour to the extra time needed to hash the CP blocks in the consensus result to compute the luck value.
Since if $N$ increases, every node must also perform more hash operations.
These result do not conform to analytical result in~\Cref{sec:cons-duration}.
Nevertheless, the difference is minor and there are ways to optimise the luck value computation.
For example, the luck value can be computed by the facilitators and are sent with the consensus result.
Then the non-facilitator nodes simply use it if there are enough signatures.

@@

Some fluctuations exist in~\Cref{fig:tx-comms-fixed}, this is due to our caching mechanism.
We send validation requests at the same rate as transactions.
Upon receiving an (remote) agreed fragment, the caching mechanism inspects all the transactions in the agreed fragment and attempts to verify as many as it can,
rather than just the transaction in the original validation request.
However, it may be the case that the agreed fragments arrive later than the validation request interval.
Then it is possible to have sent two or more validation requests for some transactions in the same local agreed fragment.
In this case, the remote would respond with two or more of the same agreed fragments,
which results in extra (wasted) communication cost, and this is the source of the fluctuation seen in~\Cref{fig:tx-comms-fixed}.
The result in~\Cref{fig:tx-comms-random} reinforces our argument.
It is a lot more stable because for every transaction it is almost always the case that an agreed fragment is needed to validate it.
% It is interesting to note that some fluctuations exist in the graphs.
% This is likely due to the way how the communication cost are recorded.
% We maintain the total amount of data sent and received in a few variables.
% On every message that is sent or received, we increment the relevent variable by the size of the message.
% However, it is infeasible to log the values every time a message is sent or received as there is simply too much data for our build server to hold.
% Therefore we only log every 5 seconds.
% This does not always align with the time when a round finishes, thus some of the data may not represent reality accurately.
%  TODO this can be fixed

@@

The difference in magnitude between \Cref{fig:global-throughput-fixed} and \Cref{fig:global-throughput-random} is caused by the caching mechanism mentioned earlier.
If a new agreed fragment needs to be transmitted to validate every transaction then it puts a toll on the network infrastructure.
The low transaction rate in~\Cref{fig:global-throughput-random} is caused by fact that the network infrastructure cannot keep up with our demand.

@@

\section{Communication cost with varying number of facilitators}
Up to this point we focused on the effect of communication cost, throughput and so on with respect to the population size.
In this section we consider a varying number of facilitators,
which gives us an insight on our system performance may perform when the number of facilitators are higher than 32.

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/consensus-communication-cost-vs-facilitators}
    \caption{Transactions are with fixed neighbours.}
    \label{fig:facilitators-comms-fixed}
  \end{subfigure}%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/consensus-communication-cost-vs-facilitators}
    \caption{Transactions are with random neighbours.}
    \label{fig:facilitators-comms-random}
  \end{subfigure}
  }
  \caption{The communication cost of the consensus protocol increase polynomially with respect to the number of facilitators as expected from the ACS communication complexity.}
  \label{fig:facilitators-comms}
\end{figure}

>>>>

To build up the confidence of the horizontal scalability claim from our analytical results,
we must verify the same property experimentally.
Hence, we give an open source prototype implementation of our system.
Our implementation runs on a network with over 1000 node.
Each node sends transactions to each other autonomously, 
which allows us to evaluate our analytical claims.

@@

The structure of the implementation is primarily made up of three modules---\texttt{acs}, \texttt{trustchain} and \texttt{node}.
\texttt{acs}, as its name suggests, implements ACS.
ACS uses erasure code in one of its sub-protocols (reliable broadcast).
Thus we use the liberasurecode\footnote{\url{https://github.com/openstack/liberasurecode}} library for its Reed-Solomon error correcting code functionality.
An implementation detail is that liberasurecode cannot create more than 32 code blocks\footnote{
  The 32 code blocks limitation is hardcoded in the source file, see
  \url{https://github.com/openstack/liberasurecode/blob/0794b31c623e4cede76d66be730719d24debcca9/include/erasurecode/erasurecode.h\#L35}
}, we discuss the effect of this in~\Cref{sec:experimental-setup}.
The \texttt{acs} module provides a small interface to the caller to start and stop the consensus process and also retrieve results.
The \texttt{trustchain} module implements the Extended TrustChain data structure.
It also provides the essential algorithm necessary to interact with Extended TrustChain such as 
$\textsf{new\_tx}(\cdot)$, $\textsf{new\_cp}(\cdot)$ and $\textsf{agreed\_fragment}(\cdot)$.
Finally, the \texttt{node} module ties everything together.
It implements the consensus protocol, the transaction protocol and the validation protocol.

@@

The goal of the experiment is to run the three protocols---consensus protocol,
transaction protocol and validation protocol---simultaneously and analyse the communication cost, the consensus duration and the global throughput.
We investigate these properties under the following four parameters.
\begin{enumerate}
  \item The transaction rate $r_{\text{tx}}$ per node. This is not comparable to the others because it is fixed at 2 TX/s.
        Nevertheless, it is a good value because, as we show later,
        it hits a bottleneck in extreme cases which helps us understand the limitations of our design.
  \item The number of facilitators $n \in \{4, 8, \dots, 32\}$.
        The maximum $n$ is 32 because the limitation in liberasurecode mentioned in \Cref{sec:implementation},
        but our results give a good indication of how our system may function for larger numbers of $n$.
  \item The population size $N \in \{200, 300, \dots, 1200\}$.
        $N$ stops at 1200 is due to our physical setup, which we describe next.
  \item The two modes of transaction.
        The first mode is the \emph{fixed-neighbour} mode where nodes only transact with their immediate neighbour which is predetermined. 
        This minimises the data volume per validated transaction because agreed fragment can be cached.
        The second mode is in the other extreme which we call the \emph{random-neighbour} mode,
        where every transaction is with a random node out of the $N$ nodes in the system.
        It is unlikely that the agreed fragments can be reused.
\end{enumerate}

The experiment is run on the DAS-5\footnote{\url{https://www.cs.vu.nl/das5/}}.
From now on, we use ``machines'' to refer to DAS-5 nodes and ``nodes'' to refer to a running instance in our system.
On DAS-5 we use up to 30 machines, for each machine we use up to 40 nodes.
This gives us the aforementioned 1200 number.
With this setup, we cannot run more nodes because the every machine only has 65535 ports available (minus the reserved ones).
But 40 nodes each need 1200 TCP connections which is 48000 TCP connections per machine and that is inching close to the limit.
While it is possible to have more TCP connections per machine,
but it requires additional network interface which is something we do not control on the DAS-5.
Nevertheless, running the system with 1200 nodes gives a good indication of its scalability properties as we show later.

To coordinate nodes on many different machines,
we employ a discovery server to inform every node the IP addresses and port numbers of every other node.
It is only run before the experiment and is not used during the experiment.

@@

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
    \begin{subfigure}[b]{.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-fixed/round-duration-vs-population}
      \caption{Transactions are with fixed neighbours.}
      \label{fig:round-duration-fixed}
    \end{subfigure}%
    \begin{subfigure}[b]{0.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-random/round-duration-vs-population}
      \caption{Transactions are with random neighbours.}
      \label{fig:round-duration-random}
    \end{subfigure}%
  }
  \caption{Round duration does not increase linearly with the population size. 
  This is likely due to the additional hashing operation required for larger consensus result.}
  \label{fig:round-duration}
\end{figure}

\Cref{fig:round-comms} shows the relationship between the communication cost of the consensus protocol per round and the population size.
The most important aspect is that these results show a linear increase.
This reinforces our analytical result in~\Cref{sec:cons-complexity}.
Note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
the magnitude of the communication cost is similar.
Both peak at about 100 MB.
This is expected because the consensus protocol is decoupled from the transaction protocol and the validation protocol.
Finally, the rate for which the communication cost increases is higher when the number of facilitators is higher.
This is also expected because the ACS algorithm has an $n^2$ term in its communication complexity.

We are also interested in how communication costs translate to time.
Hence, for the same experiment, we record the duration in seconds and the result is shown in~\Cref{fig:round-duration}.
Interestingly, the duration is not entirely linear.
We attribute this behaviour to the extra time needed to hash the CP blocks in the consensus result to compute the luck value.
Since if $N$ increases, every node must also perform more hash operations.
These results do not conform to analytical result in~\Cref{sec:cons-duration}.
Nevertheless, the difference is minor and there are ways to optimise the luck value computation.
For example, the luck value can be computed by the facilitators and are sent with the consensus result.
Then the non-facilitator nodes simply use it if there are enough signatures.

@@

Some fluctuations exist in~\Cref{fig:tx-comms-fixed}, this is due to our caching mechanism.
We send validation requests at the same rate as transactions.
Upon receiving a (remote) agreed fragment, the caching mechanism inspects all the transactions in the agreed fragment and attempts to verify as many as it can,
rather than just the transaction in the original validation request.
However, it may be the case that the agreed fragments arrive later than the validation request interval.
Then it is possible to have sent two or more validation requests for some transactions in the same local agreed fragment.
In this case, the remote would respond with two or more of the same agreed fragments,
which results in extra (wasted) communication cost, and this is the source of the fluctuation seen in~\Cref{fig:tx-comms-fixed}.
The result in~\Cref{fig:tx-comms-random} reinforces our argument.
It is a lot more stable because for every transaction it is almost always the case that an agreed fragment is needed to validate it.

@@

The difference in magnitude between \Cref{fig:global-throughput-fixed} and \Cref{fig:global-throughput-random} is caused by the caching mechanism mentioned earlier.
If a new agreed fragment needs to be transmitted to validate every transaction then it puts a toll on the network infrastructure.
The low transaction rate in~\Cref{fig:global-throughput-random} is caused by the fact that the network infrastructure cannot keep up with our demand.

@@

\section{Communication cost with varying number of facilitators}
Up to this point we focused on the effect of communication cost, throughput and so on with respect to the population size.
In this section, we consider a varying number of facilitators,
which gives us an insight into our system performance may perform when the number of facilitators is higher than 32.

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/consensus-communication-cost-vs-facilitators}
    \caption{Transactions are with fixed neighbours.}
    \label{fig:facilitators-comms-fixed}
  \end{subfigure}%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/consensus-communication-cost-vs-facilitators}
    \caption{Transactions are with random neighbours.}
    \label{fig:facilitators-comms-random}
  \end{subfigure}
  }
  \caption{The communication cost of the consensus protocol increases polynomially with respect to the number of facilitators as expected from the ACS communication complexity.}
  \label{fig:facilitators-comms}
\end{figure}

### 2017-08-06 thesis/chapters/ch6-conclusion.tex
Write abstract
Write preface and update Bitcoin situation
Spelling/grammar check for ch3, 4, 5, 6

In this work we introduced an application neutrual blockchain system which we call Checkpoint Consensus.
We add checkpoint block to the existing TrustChain data structure for use in our consensus protocol.
The round based consensus protocol uses ACS as a building block to reach consensus on checkpoint blocks.
The consensus result lets nodes elect new facilitators and create new checkpoint blocks.
Node make transactions via the transaction protocol,
which is adapted from an existing work called True Halves.
We introduce a validation protocol which ensure that if agreed fragments for some transaction exists,
then nodes reach agreement on the validity of that transaction.
The novelty of the validation protocol is that it uses point-to-point communication,
i.e. nodes only validate the transactions of interest,
this enables our horizontal scalability property.

@@

\section{Future work}
While our system has excellent scalability properties, it is not without limitations.
Much of it is from the fact that it does not have a concrete application.
We do not attempt to prevent the Sybil attack, spam or DoS because the accuracy of the defence depends on the understanding of the application.
For instance, without any application it would be impossible to distinguish between a very active node from a spammer because the content of the transaction ($m$) carry no meaning.

Our experiment is carried out in a somewhat idealised world (DAS-5).
To understand our system further, we hope to experiment on a real platform with real users such as Tribler~\cite{pouwelse2008tribler}.
For instance, the TrustChain in Tribler stores the number of bytes uploaded and downloaded between users, but it has no global consensus.
It would be useful to integrate our consensus protocol into Tribler to prevent misreporting attacks.
More importantly it gives us an opportunity to observe the characteristics of our system in the real world and how it evolves with new features or alterations.

The fault tolerance property is adaquate for permissioned systems where nodes have relatively more trust between each other.
However, it is not adaquate for permissionless systems,
which usually aim to tolerate fault that involve a minority of the network,
which can be just shy of 50\%.
For future work, we propose the following two ways to improve fault tolerance.
First, use a reputation mechanism to select facilitators rather than simply selecting them randomly so that faulty nodes are less likely to be selected.
This technique of course heavily depends on the application,
for P2P file sharing systems the reputation score may be the number of bytes contributed to the network.
Second, instead of using ACS we use a proof-of-stake (PoS) scheme which means that nodes with more stake in the system gets more vote.
In an ideal PoS scheme, the system can tolerate any fault if the majority of the stake holders are honest.

>>>>

In this work we introduced an application neutral blockchain system which we call Checo.
We add checkpoint block to the existing TrustChain data structure for use in our consensus protocol.
The round based consensus protocol uses ACS as a building block to reach consensus on checkpoint blocks.
The consensus result lets nodes elect new facilitators and create new checkpoint blocks.
Nodes make transactions via the transaction protocol,
which is adapted from an existing work called True Halves.
We introduce a validation protocol which ensures that if agreed fragments for some transaction exists,
then nodes reach agreement on the validity of that transaction.
The novelty of the validation protocol is that it uses point-to-point communication,
i.e. nodes only validate the transactions of interest,
this enables our horizontal scalability property.

@@

\section{Future work}
While our system has excellent scalability properties, it is not without limitations.
Much of it is from the fact that it does not have a concrete application.
We do not attempt to prevent the Sybil attack, spam or DoS because the accuracy of the defence depends on the understanding of the application.
For instance, without any application, it would be impossible to distinguish between a very active node from a spammer because the content of the transaction ($m$) carry no meaning.

Our experiment is carried out in a somewhat idealised world (DAS-5).
To understand our system further, we hope to experiment with a real platform with real users such as Tribler~\cite{pouwelse2008tribler}.
For instance, the TrustChain in Tribler stores the number of bytes uploaded and downloaded between users, but it has no global consensus.
It would be useful to integrate our consensus protocol into Tribler to prevent misreporting attacks.
More importantly, it gives us an opportunity to observe the characteristics of our system in the real world and how it evolves with new features or alterations.

The fault tolerance property is adequate for permissioned systems where nodes have relatively more trust between each other.
However, it is not adequate for permissionless systems,
which usually aim to tolerate fault that involve a minority of the network,
which can be just shy of 50\%.
For future work, we propose the following two ways to improve fault tolerance.
First, use a reputation mechanism to select facilitators rather than simply selecting them randomly so that faulty nodes are less likely to be selected.
This technique of course heavily depends on the application,
for P2P file sharing systems the reputation score may be the number of bytes contributed to the network.
Second, instead of using ACS we use a proof-of-stake (PoS) scheme which means that nodes with more stake in the system get more vote.
In an ideal PoS scheme, the system can tolerate any fault if the majority of the stake holders are honest.
